Titeln på avhandlingens språk

### 2022-10-21 background.tex
writing

With this in mind, we can reformulate the results this work as follows:
\begin{enumerate}
    \item System slow-down induces \emph{additional} behavioral slow-down which scales with the decrease in system responsiveness.
    Compared to the unimpaired case, participants were on average \SI{12}{\percent} slower when subject to a mean \ac{TTF} of \SI{2.475}{\second}, and a further \num{14} percentage points slower at a mean \ac{TTF} of \SI{4.5}{\second}.
    Users also become progressively slower the longer they spend in a degraded system state.

@@

    \item Humans get faster at performing steps as the task progresses.
    For sequences of \num{12} steps in an unimpaired application state, humans executed the final four steps on average \SI{36}{\percent} faster than they did the first four.
    However, this effect is dampened by reduced system responsiveness, and disappears at the highest levels of system impairment.

>>>>

With this in mind, we can reformulate the conclusions of this work as follows:
\begin{enumerate}
    \item System slow-down induces \emph{additional} behavioral slow-down which scales with the decrease in system responsiveness.
    Compared to the unimpaired case, participants were on average \SI{12}{\percent} slower when subject to a mean \ac{TTF} of \SI{2.475}{\second}, and a further \num{14} percentage points slower at a mean \ac{TTF} of \SI{4.5}{\second}.
    Users also become progressively slower the longer they spend in a degraded system state.

@@

    \item\label{item:speedup} Humans get faster at performing steps as the task progresses.
    For sequences of \num{12} steps in an unimpaired application state, humans executed the final four steps on average \SI{36}{\percent} faster than they did the first four.
    However, this effect is dampened by reduced system responsiveness, and disappears at the highest levels of system impairment.

### 2022-10-21 model.tex
writing

\subsection{Model verification}\label{ssec:model:verification}
\todo[inline]{Needs better name.}
\todo[inline]{Rework with new tweaked models.}
\todo[inline]{Needs to include first the synthetic setup, then "real" setup.}

We use Python~\num{3.10} to implement two variants of the above described timing model.
The first of these samples execution times directly from the empirical data from~\cite{olguinmunoz:impact2021}; we will refer to it hereafter as the \emph{empirical model}.
The second corresponds to one which first fits a theoretical \ac{exGaussian} distribution to the data before sampling; we will refer to this one as the \emph{theoretical model}.
We also implement three simpler reference timing models:
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item a model which returns a constant execution time, corresponding to the mean execution time of all samples associated with the lowest impairment and the shortest duration
    % \item a model which randomly samples from the raw data without any grouping
    \item a model which fits an \ac{exGaussian} distribution to the raw data, without any grouping, and then randomly samples from it  
\end{enumerate*}.
We will refer to these models as the \emph{constant} and \emph{fitted-naive} model, respectively.

\todo[inline]{Insert synthetic, controlled setup here.}

@@

We show the results of these runs in \cref{fig:ctrlttf}, and we can notice two interesting behaviors of the realistic models.
One, all realistic models are faster than the reference models at the lowest \ac{TTF}.
When compared to the fitted-naive model, the reduction in total duration was on average \textasciitilde\SI{12}{\percent} for realistic models with low neuroticism, and \textasciitilde\SI{6.8}{\percent} for the same models with high neuroticism.

@@

\todo[inline]{explain whyyyyy}

@@

These models are integrated into client software for communicating with an adapted and instrumented version of the Gabriel~\cite{Chen2015LEGO,Chen2018application} \ac{WCA} backend running variants of the LEGO task.
In the following, we use these client-server loops to illustrate the implications of the realistic timing models, which react to changes in system responsiveness in real-time, when compared to the static, naive model.
Two scenarios are studied, using the testbed shown in \cref{fig:expsetup}:
\begin{enumerate}
    \item\label{verification:scaling} A scenario to study the effects of decreasing system responsiveness due to resource and network congestion.
    An increasing number of client-server loops (\numlist{1;4;7;10} loops) are deployed in parallel for each model.
    Neuroticism is set to a constant value of \num{0.5}, and each combination of model and number of clients is repeated \num{5} times for better statistical representativeness.
    \item\label{verification:neuro} A scenario in which the effects of neuroticism are studied.
    Two different configurations of the realistic models, with neuroticism values of \num{0.0} (\emph{low}) and \num{1.0} (\emph{high}), are compared to a reference setup using the naive model.
    In all these setups, \num{9} loops are deployed in parallel, and an \emph{iperf3}\footnote{\href{https://iperf.fr/}{https://iperf.fr/}} client-server pair is deployed between the tenth Raspberry Pi client and the cloudlet.
    This \emph{iperf3} deployment starts \SI{120}{\second} after the start of the \num{9} model loops, and is configured to generate additional network load by constantly sending a stream of \SI{1460}{\byte} \ac{UDP} datagrams on the uplink; the rate at which these datagrams are sent is only limited by the \acs{CPU} of the Raspberry Pi client.

    We repeat each (model, neuroticism value) configuration --- (naive, \acs{N/A}), (empirical, low), (empirical, high), (theoretical, low), and (theoretical, high) --- \num{5} times for better statistical accuracy.
\end{enumerate}

Both scenarios described above use a variant of one of the \num{169}-step tasks used in~\cite{olguinmunoz:impact2021}, truncated to \num{45} steps.

@@

On the other, the realistic models experience a mean slowdown of about \SI{5.13}{\percent} at the highest level of resource contention (\num{10} concurrent loops) when compared to the naive model, even though only about \SI{22}{\percent} of \acp{TTF} in this scenario reach the medium impairment level and practically no \acp{TTF} pass threshold for the highest level of impairment.
This showcases that the additional user slow-down due to reduced system responsiveness, originally described in \textcite{olguinmunoz:impact2021}, causes a noticeable and significant extension of system utilization time, even at lower and medium levels of impairment.
\todo[inline]{Add another setup with additional load and discuss}

>>>>

\subsection{Model verification}\label{ssec:model:verification}
\todo[inline]{Needs better name?}

We use Python~\num{3.10} to implement two variants of the above described timing model.
The first of these samples execution times directly from the empirical data from~\cite{olguinmunoz:impact2021}; we will refer to it hereafter as the \emph{empirical model}.
The second corresponds to one which first fits a theoretical \ac{exGaussian} distribution to the data before sampling; we will refer to this one as the \emph{theoretical model}.
The \ac{exGaussian} distribution was chosen due to the vast body of research supporting its suitability for the modeling of the timing of human actions~\cite{Rohrer1994analysis,Palmer2011shapes,Marmolejo2022generalised}.
Collectively, we will refer to both these models interchangeably as the \emph{realistic models} or \emph{second-order models}.
\todo[inline]{Explain why using ExGaussian?}

We also implement two simpler reference timing models:
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    % \item a model which returns a constant execution time, corresponding to the mean execution time of all samples associated with the lowest impairment and the shortest duration
    \item a model which randomly samples from the raw data without any grouping
    \item a model which fits an \ac{exGaussian} distribution to the raw data, without any grouping, and then randomly samples from it  
\end{enumerate*}.
We will refer to these models individually as the \emph{naive} and \emph{fitted-naive} model, respectively, and collectively as the \emph{firt-order models}.

@@

We show the results of these runs in \cref{fig:ctrlttf}.
Two interesting behaviors of the realistic models can be noticed in this plot.
One, all realistic models are faster than the reference models at the lowest \ac{TTF}.
When compared to the fitted-naive model, the reduction in total duration was on average \textasciitilde\SI{12}{\percent} for realistic models with low neuroticism, and \textasciitilde\SI{6.8}{\percent} for the same models with high neuroticism.

@@

These behaviors clearly stem from the effects identified in \textcite{olguinmunoz:impact2021}, described in \cref{ssec:plos}.
At low levels of impairments, i.e.\ small \acp{TTF}, the speed-up effect dominates, bringing execution times to a minimum and consequently resulting in shorter total durations.
On the other hand, at high levels of impairment, the slow-down (pacing) effect starts to take over, pushing step execution times and total durations up.
However, only in the high neuroticism case do execution times eventually surpass the reference model, which is in line with our findings with respect to the modulating effects of neuroticism.

@@

These models are then integrated into client software for communicating with an adapted and instrumented version of the Gabriel~\cite{Chen2015LEGO,Chen2018application} \ac{WCA} backend running variants of the LEGO task.
In the following we use these client-server loops together with the testbed show in \cref{fig:expsetup} to illustrate the implications of the realistic timing models.
Our goal is to understand the consequences of taking second order effects into consideration when modeling execution times, versus simply relying on first-order statistics.

\todo[inline]{Update with new deployment, 802.11b/g wifi.}

We deploy an increasing number of client-server loops (\numlist{1;4;7;10} loops) in parallel on the testbed and record the resulting timing metrics.
These setups are repeated \num{5} times per model for better statistical representativeness; realistic models parameterized with \emph{low} (\num{0.0}) and \emph{high} (\num{1.0}) neuroticism are considered separate models for the purpose of this experiment.
The task used for all repetitions corresponds to a variant of the \num{169}-step tasks used in~\cite{olguinmunoz:impact2021}, truncated to \num{45} steps plus a special initial trigger step.

@@

On the other, the realistic models experience a mean slowdown of about \SI{5.13}{\percent} at the highest level of resource contention (\num{10} concurrent loops) when compared to the naive model, even though only about \SI{22}{\percent} of \acp{TTF} in this scenario reach the medium impairment level and practically no \acp{TTF} pass threshold for the highest level of impairment.
This showcases that the additional user slow-down due to reduced system responsiveness, originally described in \textcite{olguinmunoz:impact2021}, causes a noticeable and significant extension of system utilization time, even at lower and medium levels of impairment.
